Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-04-03 Manuscript/manuscript.tex
Removing TODOs from MS and also from SI

Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
Because POPS carries a physical total charge as ions in water, we apply the theoretically derived scaling factor for
ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. The lipids with a total charge of -0.75 also neutralize
the counterion charges of +0.75 in simulations with ECC-ions \cite{kohagen16}.
Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
Further optimization of parameters was not done in this work.
\todo{Add here a reference to the parameters.}
% we used exactly the pre-determined scaling factor 
%to represent the effects of electronic polarization
%yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
%polar parts of the phospholipid,
%i.e., charges on all atoms except on alkyl tails, 
%which do not encounter direct contacts with ions.

@@

\subsection{Measurements of acyl chain order parameters}
A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
parameters
\begin{equation}\label{OPequation}
  S_{\rm{CH}}=\frac{1}{2}\langle3\cos^2\theta-1\rangle,
\end{equation}
  where $\theta$ denotes the angle of the C--H bond
with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
(using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
therefore a length of 10.79 us which correspond to a nutation frequency of 46.35 kHz. Increments of 18x2 recoupling
pulses where used giving a spectral width of 2.6 kHz in the indirect dimension and a total number of 32 points in
the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} \todo{Is this reference to the correct figure?})
is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
\todo{ I could not find this reference [D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]}
was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
time of 0.1 s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50 kHz, and with dwell time giving a spectral
width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
described e.g. in Ref. \citenum{ferreira08}. The dipolar splittings in the crowded spectral region
between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
and the POPS order parameters calculated from simulations in this work.

@@

The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}
\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{NMRlipidsIV}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
  \todo{We should use the same color coding for the models as in ECC-POPC paper, now it is opposite.} \\
  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
\end{figure}

@@

\begin{table}[tb!] 
\centering
  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & APL / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
%    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
%    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
%    \hline 
  \end{tabular} \\
  \todo{I removed the value from Lipid17/ff99 simulation because this was not listed in Table I
    and I think that it is more clear if we refer only to the NMRlipids IV result in the text.
    I also changed the experimental value to correspond the value in the cited paper and added the error bar.}
\end{table}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.
  SAMULI: I think that these are not promised in the current text and addition is not necessary.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?
  SAMULI: I feel now that we should keep the figure as it is. Also the POPC:POPS (5:1) results are now referred in the text.
  However, if something is moved to SI, it should be POPS in main text, mixture in SI.}
\end{figure} 
Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters \cite{catte16,NMRlipidsIV},
which is the case also here as the sodium counterion binding affinity to
POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
mixtures\cite{roux90,NMRlipidsIV}.
Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
to lipid bilayer in simulations containing PS using two different approaches.

@@

% This§: 
% adsorption and affinities of Na and K to PC:PS memb
% from densities and rel.surf.excess
% 
%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%
The influence of negatively charged PS on the counterion binding affinity can
be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01$ nm$^{-2}$ for 1 M sodium in pure ECC-POPC
simulation was reported in the previous work \cite{melcr18}.
The presence of $\sim$17\% POPS in POPC bilayer increases the
%relative surface excess of sodium with respect to water from negative
value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 nm$^{-2}$ 
for 0.621 M sodium, while value for potassium remains negative,
$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 nm$^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
ECC-lipid simulations. The response of headgroup order parameters to the addition of NaCl are
slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
increase in binding affinity due to PS lipids would not be affected.
%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
%showing that the density profile of \ce{K^+} along the membrane normal 
%decays clearly faster towards the membrane centre compared to the density of water. 
%In contrast, the density profile of \ce{Na^+} has a shoulder 
%at the region of the phosphate groups,
%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
%by yielding a small positive value of $\Gamma^{w}_{Na}$.

@@

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
  \caption{\label{fig:cacl_complexes} 
    Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
    Analysis was done for both systems by considering all lipids (blue and violet) and
    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange).
    %All lipids were taken into account with the exception of the complexes in light green and orange, 
    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
    Clusters of four or more lipids were not observed in either membrane.
    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from ?? Add Zenodo reference, if available. 
  }
  \todo{Add definition used for the coordination} \\
  \todo{Add link to the Zenodo, if available.}
\end{figure}

>>>>

Here, we apply the ECC to the Amber based lipid17 parameters \cite{lipid17-future} of POPS lipid with a monovalent negative charge.
Because POPS carries a physical total charge as ions in water, we apply the theoretically derived scaling factor for
ions $f_q = 0.75$ \cite{leontyev09} to the partial charges of POPS. The lipids with a total charge of -0.75 also neutralize
the counterion charges of +0.75 in simulations with ECC-ions \cite{kohagen16}.
Following our previous work for POPC, we use the scaling factor of $f_\sigma = 0.89$ for the Lennard-Jones~$\sigma$ parameters and
scale the parameters of atoms only in the headgroup, glycerol backbone, and carbonyl regions \cite{melcr18}.
Further optimization of parameters was not done in this work. \cite{ECC-POPC_nacl_cacl2_files}
% we used exactly the pre-determined scaling factor 
%to represent the effects of electronic polarization
%yielding a total charge of $-0.75$ in the ECC formalism. \cite{martinek17, leontyev09, melcr18}
Similarly to the ECC-POPC model \cite{melcr18}, we applied the ECC only to the
non-tail atoms, i.e., to atoms in the headgroup, glycerol backbone, and carbonyl regions. 
%polar parts of the phospholipid,
%i.e., charges on all atoms except on alkyl tails, 
%which do not encounter direct contacts with ions.

@@

\subsection{Measurements of acyl chain order parameters}
A R-type Proton Detected Local Field (R-PDLF) experiment was performed to determine C--H bond order
parameters
\begin{equation}\label{OPequation}
  S_{\rm{CH}}=\frac{1}{2}\langle3\cos^2\theta-1\rangle,
\end{equation}
  where $\theta$ denotes the angle of the C--H bond
with the bilayer normal and the angular brackets define a time average on a time scale of approximately 1 microsecond.
The experiment was done using a Bruker Avance III 400 spectrometer operating at a 1H Larmor frequency of 400.03 MHz
equipped with a standard 4 mm CP-MAS HXY probe. The set up of the R-PDLF experiment was the following
(using the notation from figures 1c and 2c in the original publication describing the R-PDLF experiment \cite{dvinskikh04}).
The magic angle spinning (MAS) frequency used was 5.15 kHz. The recoupling pulses in the R18 blocks had
therefore a length of 10.79 us which correspond to a nutation frequency of 46.35 kHz. Increments of 18x2 recoupling
pulses where used giving a spectral width of 2.6 kHz in the indirect dimension and a total number of 32 points in
the indirect dimension were recorded. These settings enable to record dipolar slices with a C--H bond order parameter
resolution of +- 0.01. The C--H bond order parameter determined from a given dipolar
splitting, $\Delta\nu$ (see e.g. Fig. \ref{R-PDLFslices} in SI)
is equal to $\Delta\nu/(0.315\times21.5 kHz)$, where 0.315 is
the scaling factor of the R18 recoupling sequence and 21.5 kHz is the maximum $^1$H-$^{13}$C dipolar coupling for a C--H bond.
The refocused-INEPT transfer \cite{morris79,burum80} %[G. Morris and R. Freeman, J. Am. Chem. Soc. 101, 760 (1979);
\todo{ I could not find this reference [D. Burum, R. Ernst, J. Mag. Res. 39, 163 (1980)]}
was used for transferring the polarization from $^1$H nuclei to the covalently bond $^{13}$C nuclei with the delays $\tau_1$ = 1.94 ms and $\tau_2$ = 0.97 ms
(set as multiples of the MAS rotation period). The pulses for the refocused-INEPT sequence had a nutation frequency equal to 63.45 kHz.
For acquiring the spectra, a total number of 1024 transients were recorded for each point in the indirect dimension, using an acquisition
time of 0.1 s under SPINAL64 $^1$H decoupling \cite{fung00} with a nutation frequency of 50 kHz, and with dwell time giving a spectral
width of 200 ppm for the $^{13}$C chemical shift dimension. The set of free induction decays recorded were then processed as
described e.g. in Ref. \citenum{ferreira08}. The dipolar splittings in the crowded spectral region
between 29 and 30 ppm were assigned based on the previous assignments for POPC \cite{ferreira13} 
and the POPS order parameters calculated from simulations in this work.

@@

The X-ray scattering form factors and C-H bond order parameters from NMR experiments are good measures to validate
the lipid bilayer structure in MD simulations against experiments because they
are related to the bilayer dimensions (area per molecule and thickness) and
conformational fluctuations of individual lipids, respectively, and 
can be directly calculated from MD simulations~\cite{ollila16}.
The X-ray scattering form factors and C-H bond order parameters in
the headgroup and glycerol backbone region of POPS lipid bilayer are already available in the literature \cite{kucerka14,NMRlipidsIV}.
Here, we measure also the C--H bond order parameters of the acyl chain region using natural abundance $^{13}$C NMR.
We first determined the order parameters of each peak corresponding different C--H bonds detected
in the first dimension (Fig. \ref{R-PDLF}) from the dipolar slices in the indirect dimension (Fig. \ref{R-PDLFslices}).
Because the exact assignment of these peaks was not possible, we assigned the
acyl chain order parameters to mimic the profile predicted by the MD simulations (Fig. \ref{simVSexpNOions_POPS}).
\todo{Tiago has to check and update this and figures in SI.}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/ecc_pops/Order-parameters_form-factors_exp-L17-ECC-lipids.pdf}
  \includegraphics[width=\figwidth]{../img/ecc_pops/pops_chemfig.pdf} 
\hfill
  \caption{\label{simVSexpNOions_POPS} 
    \textbf{Top:} X-ray scattering form factors from simulations with Lipid17/Dang \citep{lipid17-future, dang2006} and 
    the ECC-POPS/ECC-ions \cite{martinek17, Pluhackova2016} compared with experiments~\citep{kucerka14} at 298~K. 
    \textbf{Middle:} Order parameters of POPS head group, glycerol backbone and acyl chains  
    from simulations with the Lipid17 \citep{lipid17-future} and the ECC-POPS models 
    compared with experiments at 298~K. \citep{NMRlipidsIV}
    Open/closed symbols are used for palmitoyl/oleoyl chains of POPS. 
    \textbf{Bottom:} The chemical structure of POPS and the labeling of the carbon segments. 
  }  
  \todo{J: flip the chemical structure so that it agrees with the left-to-right reading of the plot.} \\
  \todo{The experimental order parameters from sn-1 chain are missing from the figure.} \\
  \todo{Inverted axis and positive value as label would be more clear for the order parameter figure.}
\end{figure}

@@

\begin{table}[tb!] 
\centering
  \caption{Area per lipid (APL) of POPS bilayers from simulations with different force fields at 298~K with \ce{Na^+} counterions. \label{tab:apls} } 
  \begin{tabular}{l|c } 
    \multicolumn{2}{c}{POPS} \\
    model          & APL / Å$^2$    \\ 
    \hline 
    Lipid17/Dang              & 53.5$\pm$ 1.7   \\ 
%    Lipid17/ff99              & 57.9$\pm$ 1.7   \\ 
%    \hline 
    ECC-POPS/ECC-ions         & 59.8$\pm$ 1.8   \\ 
%    \hline 
    experiment (SDP model) \citep{kucerka14} & 62.7$\pm$ 1.3  \\ 
%    \hline 
  \end{tabular} \\
\end{table}

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}} and \ce{Na^{+}} counterions along the membrane normal axis
    in ECC-lipid (solid lines) and Lipid17/Dang (dashed lines) simulations of pure POPS (left) or POPC:POPS (5:1, right) bilayers.  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\end{figure} 
Binding affinity of monovalent ions to lipid bilayers depends strongly on
force field parameters \cite{catte16,NMRlipidsIV},
which is the case also here as the sodium counterion binding affinity to
POPS lipid bilayer is stronger in Lipid17/ Dang simulation than in
ECC-POPS simulation (Fig. \ref{fig:POPS-counterions-dens}).
The cation binding affinities to lipid bilayers in simulations have been previously evaluated
against experiments using the order parameters of $\alpha$ and $\beta$ carbons in the
phosphatidylcholine (PC) lipid headgroup (see Fig.~\ref{fig:chemstruct_pc_ps} for the labeling) \cite{catte16,melcr18,NMRlipidsIV},
which decrease proportionally to the bound charge according to the ''electrometer concept'' \citep{seelig87}.
Althought the response of PS lipid headgroup order parameters to the bound charge is also systematic,
it is less well undestood and the addition of cations may cause phase transitions in negatively charged
bilayers \cite{feigenson86,mattai89,roux91,roux90}. Therefore, the cation binding affinity to bilayers with
PS lipids has been evaluated measuring the PC headgroup order parameters from POPC:POPS (5:1)
mixtures\cite{roux90,NMRlipidsIV}.
Following our previous work \cite{NMRlipidsIV}, we evaluate the counterion binding affinity
to lipid bilayer in simulations containing PS using two different approaches.

@@

% This§: 
% adsorption and affinities of Na and K to PC:PS memb
% from densities and rel.surf.excess
% 
%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%
The influence of negatively charged PS on the counterion binding affinity can
be estimated by comparing the relative surface excesses with respect to water, $\Gamma ^{w} _{\rm ion}$,
between POPC:POPS (5:1) mixture and pure POPC \cite{melcr18}.
%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
The value of $\Gamma ^{w} _{\rm Na}=-0.11 \pm 0.01 \mathrm{nm}^{-2}$ for 1 M sodium in pure ECC-POPC
simulation was reported in the previous work \cite{melcr18}.
The presence of $\sim$17\$ POPS in POPC bilayer increases the
%relative surface excess of sodium with respect to water from negative
value to $\Gamma ^{w} _{\rm Na}=0.092 \pm 0.005 \mathrm{nm}^{-2}$ 
for 0.621 M sodium, while value for potassium remains negative,
$\Gamma ^{w} _{\rm K}=-0.123 \pm 0.005 \mathrm{nm}^{-2}$ in POPC:POPS (5:1) mixture (Fig. \ref{fig:nacl-dens_PCPS}).
As discussed above, the sodium binding to POPC:POPS (5:1) mixture may be slightly overestimated also in
ECC-lipid simulations. The response of headgroup order parameters to the addition of NaCl are
slightly overestimated also in ECC-POPC simulations (Fig 3 in Ref. \citenum{melcr18}),
suggesting that similar small overestimation of sodium binding may also occur to a pure POPC bilayer.
Therefore, the overestimation may arise from the sodium interactions with PC headgroup and
increase in binding affinity due to PS lipids would not be affected.
%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
%showing that the density profile of \ce{K^+} along the membrane normal 
%decays clearly faster towards the membrane centre compared to the density of water. 
%In contrast, the density profile of \ce{Na^+} has a shoulder 
%at the region of the phosphate groups,
%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
%by yielding a small positive value of $\Gamma^{w}_{Na}$.

@@

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../img/stoichiometry_CaCl2_comparison_Ecc-lipids_PC-vs-PCPS.pdf} \\ 
  \caption{\label{fig:cacl_complexes} 
    Relative probabilities of \ce{Ca^{2+}} ions to coordinate with a certain number of lipids
    in pure POPC bilayer with 350 mM CaCl$_2$ and in POPC:POPS (5:1) mixture with 400 mM CaCl$_2$.  
    Analysis was done for both systems by considering all lipids (blue and violet) and
    for POPC:POPS (5:1) mixture also by considering POPC and POPS lipids separately (green and orange).
    %All lipids were taken into account with the exception of the complexes in light green and orange, 
    %for which we counted only contacts with POPC resp POPS from the mixed 5\,PC:1\,PS negatively charged bilayer. 
    %The calculated probabilities of the calcium-lipid complexes also reflect only POPC (light green) resp POPS (orange). 
    % Probabilities were taken from simulations with comparable bulk concentrations of calcium around 250~mM. 
    Clusters of four or more lipids were not observed in either membrane.
    The threshold for counting coordinated lipids in a complex with \ce{Ca^{2+}} was set to $0.3\,\mathrm{nm}$, 
    as the distance between the cation and the oxygen atoms of the lipids. 
    Previously published simulation data \cite{melcr18} for pure POPC bilayers were taken directly from \cite{ECC-POPC_nacl_cacl2_files}. 
  }
\end{figure}
